---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# libproj

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
![R-CMD-check](https://github.com/paleolimbot/libproj/workflows/R-CMD-check/badge.svg)
<!-- badges: end -->

The goal of libproj is to provide access to the [PROJ](https://proj.org) C API for high-performance geometry operations within the R package framework. This package contains a copy of the PROJ library, modified slightly to eliminate errors on all the platforms checked by CRAN and GitHub actions. This means you don't have to install anything (other than the package) to take advantage of PROJ functions in R. It also means you don't need a configure script if you're writing a package that needs PROJ functionality (just `LinkingTo: libproj`).

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("paleolimbot/libproj")
```
## Example

This package only exists for its exported C API, for which headers are provided to make calling these functions from Rcpp or another package as easy and as safe as possible.

```{Rcpp}
#include <Rcpp.h>
// Packages will also need LinkingTo: libproj
// [[Rcpp::depends(libproj)]]
// needed in every file that uses proj_*() functions
#include "libproj.h"

// needed exactly once in your package or Rcpp script
// contains all the function pointers and the
// implementation of the function to initialize them
// (`libproj_init_api()`)
#include "libproj.c"

// this function needs to be called once before any proj_*() functions
// are called (e.g., in .onLoad() for your package)
// also need to specify the path to the database

using namespace Rcpp;
// [[Rcpp::export]]
void cpp_libproj_init_api() {
  libproj_init_api();
  Function systemFile("system.file");
  CharacterVector dbPath = systemFile("proj.db", _["package"] = "libproj");
  const char* dbPath0 = dbPath[0];
  proj_context_set_database_path(PJ_DEFAULT_CTX, dbPath0, NULL, NULL);
}

// regular C or C++ code that uses proj_()* functions!
// [[Rcpp::export]]
List proj_coords(List xy, std::string from, std::string to) {
  NumericVector x = xy[0];
  NumericVector y = xy[1];
  
  PJ_CONTEXT* context = PJ_DEFAULT_CTX;
  PJ* trans = proj_create_crs_to_crs(context, from.c_str(), to.c_str(), NULL);
  if (trans == NULL) {
    int errorCode = proj_context_errno(context);
    std::stringstream err;
    err << "Error creating transform: " << proj_errno_string(errorCode);
    stop(err.str());
  }
  
  NumericVector xout = clone(x);
  NumericVector yout = clone(y);
  size_t stride = sizeof(double);
  
  proj_trans_generic(
    trans, PJ_FWD,
    &(xout[0]), stride, xout.size(),
    &(yout[0]), stride, yout.size(),
    nullptr, stride, 0,
    nullptr, stride, 0
  );
  
  int errorCode = proj_errno(trans);
  proj_destroy(trans);
  
  if (errorCode != 0) {
    std::stringstream err;
    err << "Error transforming coords: " << proj_errno_string(errorCode);
    stop(err.str());
  }
  
  return List::create(_["x"] = xout, _["y"]  = yout);
}
```

```{r}
cpp_libproj_init_api()

sf::st_transform(sf::st_sfc(sf::st_point(c(-64, 45)), crs = 4326), 26920)

# note that the PROJ default axis ordering (at the C level)
# is not what you expect!
proj_coords(list(45, -64), "EPSG:4326", "EPSG:26920")
```
